Postural developments modulate children's visual access to social information

### 2020-04-30 writeup-paper/draft_apa_march2020.Rmd
edits to paper draft
make figures plausibly work in-line
add marginal r^2 analysis
other typos and edits

abstract: |
  The ability to process social information is a critical component of children's early linguistic and cognitive development.
  However, as children reach their first birthday, they begin to locomote themselves, dramatically changing the way they see the world around them. 
  How do these postural and locomotor developments affect children's access to the social information relevant for word-learning? 
  We explored this question by using head-mounted cameras to record the egocentric visual perspective of 36 infants' (at 8, 12, and 16 months of age) during a naturalistic play session with their caregiver. To estimate the proportion of faces and hands in the infant view, we used a computer vision algorithm to detect the presence of faces and hands in the entire dataset.
  We found that infants\\' posture and orientation to their caregiver changed dramatically across this age range and modulated their access to this social information; infants who were sitting or standing with their caregiver at a close distance tended to have the most faces/hands in their visual field. 
  We also applied our automated analysis to the egocentric video data from a recent paper documenting how posture changes one-year-olds visual access to faces (Franchak et al., 2017), finding convergence across both methodologies and dataset and confirming previous work that motoric developments play a significant role in the emergence of children’s linguistic and social capacities.
  We suggest that the combined use of head-mounted cameras and the application of new computer vision techniques is a promising avenue for understanding the statistics of infants\\' visual and linguistic experience as they change over development.

keywords          : "Postural developments modulate children's visual access to social information"
wordcount         : "X"

@@

# Methods
### Participants
```{r echo=FALSE, include=FALSE, demographics}
all_subs <- read.csv("../data/demographics/all_participants.csv") %>%
  rename(subid = subID)
inc_subs <- read.csv("../data/demographics/demographics.csv")
all_subs$inc <- all_subs$reason=="included"

@@

### Head-mounted camera
```{r headcam, fig.env = "figure", fig.height=3, fig.pos = "H", fig.label="headcam", fig.cap = "Vertical field of view for two different headcam configurations (we used the lower in our current study)."}

@@

### Procedure
All parents signed consent documents while children were fitted with the headcam. If the child was uninterested in wearing the headcam or tried to take it off, the experimenter presented engaging toys to try to draw the child's focus away from the headcam. When the child was comfortable wearing the headcam, the child and caregiver were shown to a playroom for the free-play session. Parents were shown a box containing three pairs of novel and familiar objects (e.g., a ball and a microfiber duster, named a "zem"), and were instructed to play with the object pairs with their child one at a time, "as they typically would." All parents confirmed that their child had not previously seen the novel toys and were instructed to use the novel labels to refer to the toys. The experimenter then left the playroom for approximately 15 minutes, during which a tripod-mounted camera in the corner of the room recorded the session and the headcam captured video from the child's perspective.

### Data Processing and Annotation
Headcam videos were trimmed such that they excluded the instruction phase when the experimenter was in the room and were automatically synchronized with the tripod-mounted videos using FinalCut Pro Software. These sessions yielded 
`r round(sum(inc_subs$len.min), digits = 0)` minutes (almost a million frames) of video, with an average video length of
`r round(mean(inc_subs$len.min), digits = 2)` minutes
(min = `r round(min(inc_subs$len.min), digits = 2)`, max = `r round(max(inc_subs$len.min), digits = 2)`).

@@

\begin{figure*}
\centering
\includegraphics[width=5.5in]{images/detector_samples_banner.pdf}
\caption{\label{fig:frames} Example face and pose detections made by OpenPose (top row) and MTCNN (bottom row) from a child in each age group. The last column features a false positive from OpenPose and a false negative from MTCNN.}
\end{figure*}

@@

These trends were quantified using two generalized linear mixed-effect models estimating the proportion of faces and hands that were in view, with orientation, posture, their interaction, and scaled participant's age as fixed effects, and with random slopes for infants' orientation and posture. A summary of the coefficients of the models can be found in Table 2, confirming that infants who were sitting/standing and in front of their caregivers saw the most faces and hands. When modeling the proportion of faces seen, age was not a significant predictor; however, age remained a significant predictor when modeling the proportion of hands seen by infants. Overall, these results suggest that infants' visual access to social information is largely modulated by their posture and orientation to their caregiver, which is in turn a function of their general locomotor development. Nonetheless, these results also suggest that infants may still increase their attention to hands throughout this age range as they continue to learn about objects, their functions, and their names.

@@

```{r runModels, eval=FALSE, include=FALSE}
#wrangle data
## Using MTCNN detections for GLM since it was the best
dGLM_faces <- d %>%
  group_by(subid, posture, orientation) %>%
  summarize(countFaces=sum(faceOP==TRUE), countNotFaces = sum(faceOP==FALSE), age.at.test=age.at.test[1]) %>%
  ungroup()

dGLM_wrists <- d %>%
  group_by(subid, posture, orientation) %>%
  summarize(countWrists=sum(wristOP==TRUE), countNotWrists = sum(wristOP==FALSE), age.at.test=age.at.test[1]) %>%
  ungroup()

## run models 
faceGLM <- glmer(cbind(countFaces,countNotFaces) ~ posture * orientation + scale(age.at.test) + 
               (posture + orientation | subid), 
             dGLM_faces, family = "binomial", control=glmerControl(optCtrl=list(maxfun=20000),optimizer=c("bobyqa")))

wristGLM <- glmer(cbind(countWrists,countNotWrists) ~ orientation * posture + scale(age.at.test) + 
               (posture + orientation | subid), 
             dGLM_wrists, family = "binomial", control=glmerControl(optCtrl=list(maxfun=20000),optimizer=c("bobyqa")))

## output ---- 
cleanedNames_Int = c("Intercept","Sit","Stand","Close","Far","Age (Scaled)", "Sit*Close", "Stand*Close","Sit*Far","Stand*Far")

# 
faceGLMOut=(summary(faceGLM))
rownames(faceGLMOut$coefficients)<-cleanedNames_Int
# 
wristGLMOut=(summary(wristGLM))
rownames(wristGLMOut$coefficients)<-cleanedNames_Int
#

xtable(faceGLMOut$coefficients, digits=c(2,2,2,2,3),"Model coefficients from a generalized linear mixed models predicting the proportion of faces seen by infants.")

@@

This dataset differed from the present in two key ways that could present a challenge for our automated methodology. First, the environment that infants were immersed in with their caregivers (and experimenters)  was much larger and more varied than the play room used in the present dataset, containing multiple structures and toys in different parts of the room for infants to explore, unlike the  room used for the present dataset which was relatively small (approximately 10' x 10'). Thus, our automated methodology could fail to generalize to scenes from these more complex environments, where detecting faces and hands could arguably be a much harder task. Second, @franchak2017see used a head-mounted eye-tracker, finding that children's posture affected where children were looking within their visual field. Nonetheless, if children are often orienting their heads towards where they are allocating their attention [@yoshida2008] then we should expect to find the same pattern of results only when analyzing the information in view.

```{r loadFranchakData}
# loads pos_detections used below for franchak dataset
load(file = '../data/franchak_dataset/preprocessed/franchak_dataset_detections.RData')
```

>>>>

abstract: |
  The ability to process social information is a critical component of children's early linguistic and cognitive development.
  However, as children reach their first birthday, they begin to locomote themselves, dramatically changing the way they see the world around them. 
  How do these postural and locomotor developments affect children's access to the social information relevant for word-learning? 
  We explored this question by using head-mounted cameras to record the egocentric visual perspective of 36 infants (at 8, 12, and 16 months of age) during a naturalistic play session with their caregiver. To estimate the proportion of faces and hands in the infant view, we used a computer vision algorithm to detect the presence of faces and hands in the entire dataset.
  We found that infants\' posture and orientation to their caregiver changed dramatically across this age range and modulated their access to this social information; infants who were sitting or standing with their caregiver at a close distance tended to have the most faces/hands in their visual field. 
  We also applied our automated analysis to the egocentric video data from a recent paper documenting how posture changes one-year-olds visual access to faces (Franchak et al., 2017), finding convergence across both methodologies and dataset and confirming previous work that motoric developments play a significant role in the emergence of children’s linguistic and social capacities.
  We suggest that the combined use of head-mounted cameras and the application of new computer vision techniques is a promising avenue for understanding the statistics of infants\' visual and linguistic experience as they change over development.

keywords          : "Postural developments modulate children's visual access to social information"
wordcount         : "4295"

@@

# Methods
## Participants
```{r echo=FALSE, include=FALSE, demographics}
all_subs <- read.csv("../data/demographics/all_participants.csv") %>%
  rename(subid = subID)
inc_subs <- read.csv("../data/demographics/demographics.csv")
all_subs$inc <- all_subs$reason=="included"

@@

## Head-mounted camera
```{r headcam, fig.env = "figure", fig.height=3, fig.pos = "H", fig.label="headcam", fig.cap = "Vertical field of view for two different headcam configurations (we used the lower in our current study)."}

@@

## Procedure
All parents signed consent documents while children were fitted with the headcam. If the child was uninterested in wearing the headcam or tried to take it off, the experimenter presented engaging toys to try to draw the child's focus away from the headcam. When the child was comfortable wearing the headcam, the child and caregiver were shown to a playroom for the free-play session. Parents were shown a box containing three pairs of novel and familiar objects (e.g., a ball and a microfiber duster, named a "zem"), and were instructed to play with the object pairs with their child one at a time, "as they typically would." All parents confirmed that their child had not previously seen the novel toys and were instructed to use the novel labels to refer to the toys. The experimenter then left the playroom for approximately 15 minutes, during which a tripod-mounted camera in the corner of the room recorded the session and the headcam captured video from the child's perspective.

## Data Processing and Annotation
Headcam videos were trimmed such that they excluded the instruction phase when the experimenter was in the room and were automatically synchronized with the tripod-mounted videos using FinalCut Pro Software. These sessions yielded 
`r round(sum(inc_subs$len.min), digits = 0)` minutes (almost a million frames) of video, with an average video length of
`r round(mean(inc_subs$len.min), digits = 2)` minutes
(min = `r round(min(inc_subs$len.min), digits = 2)`, max = `r round(max(inc_subs$len.min), digits = 2)`).

@@

```{r frames, fig.env = "figure", fig.height=3, fig.pos = "H", fig.label="headcam", fig.cap = "Example face and pose detections made by OpenPose (top row) and MTCNN (bottom row) from a child in each age group. The last column features a false positive from OpenPose and a false negative from MTCNN."}

img <- png::readPNG("images/detector_samples_banner.png")
grid::grid.raster(img)
```

@@

```{r runModels, eval=TRUE, include=TRUE}
library(MuMIn)

#wrangle data
## Using MTCNN detections for GLM since it was the best
dGLM_faces <- d %>%
  mutate(subid = as.factor(subid)) %>%
  group_by(subid, posture, orientation) %>%
  summarize(countFaces=sum(faceOP==TRUE), countNotFaces = sum(faceOP==FALSE), age.at.test=age.at.test[1]) %>%
  ungroup()

dGLM_wrists <- d %>%
  mutate(subid = as.factor(subid)) %>%
  group_by(subid, posture, orientation) %>%
  summarize(countWrists=sum(wristOP==TRUE), countNotWrists = sum(wristOP==FALSE), age.at.test=age.at.test[1]) %>%
  ungroup()

### FACES

# Age as fixed effect, full random effects
faceGLM_age_only <- glmer(cbind(countFaces,countNotFaces) ~  scale(age.at.test) +
(posture + orientation | subid), dGLM_faces, family = "binomial", control=glmerControl(optCtrl=list(maxfun=20000),optimizer=c("bobyqa")))

# Age and posture and orientation as fixed effects, full random effects
faceGLM_no_int <- glmer(cbind(countFaces,countNotFaces) ~ posture + orientation + scale(age.at.test) +  (posture + orientation | subid), dGLM_faces, family = "binomial", control=glmerControl(optCtrl=list(maxfun=20000),optimizer=c("bobyqa")))

# Age and posture * orientation as fixed effects, full random effects
faceGLM_full <- glmer(cbind(countFaces,countNotFaces) ~ posture * orientation + scale(age.at.test) +  (posture + orientation | subid), dGLM_faces, family = "binomial", control=glmerControl(optCtrl=list(maxfun=20000),optimizer=c("bobyqa")))

### WRISTS
# Age as fixed effect, full random effects
wristGLM_age_only <- glmer(cbind(countWrists,countNotWrists) ~ scale(age.at.test) +  (posture + orientation | subid), dGLM_wrists, family = "binomial", control=glmerControl(optCtrl=list(maxfun=20000),optimizer=c("bobyqa")))

wristGLM_no_int <- glmer(cbind(countWrists,countNotWrists) ~ orientation + posture + scale(age.at.test) +  (posture + orientation | subid), dGLM_wrists, family = "binomial", control=glmerControl(optCtrl=list(maxfun=20000),optimizer=c("bobyqa")))

# Age and posture and orientation as fixed effects, full random effects
wristGLM_full <- glmer(cbind(countWrists,countNotWrists) ~ orientation * posture + scale(age.at.test) +  (posture + orientation | subid), dGLM_wrists, family = "binomial", control=glmerControl(optCtrl=list(maxfun=20000),optimizer=c("bobyqa")))
```

```{r eval=FALSE}
## output
cleanedNames_Int = c("Intercept","Sit","Stand","Close","Far","Age (Scaled)", "Sit*Close", "Stand*Close","Sit*Far","Stand*Far")

#
faceGLMOut=(summary(faceGLM_full))
rownames(faceGLMOut$coefficients)<-cleanedNames_Int
#
wristGLMOut=(summary(wristGLM_full))
rownames(wristGLMOut$coefficients)<-cleanedNames_Int
#

xtable(faceGLMOut$coefficients, digits=c(2,2,2,2,3),"Model coefficients from a generalized linear mixed models predicting the proportion of faces seen by infants.")
```

```{r}
wrists_age_only = r.squaredGLMM(wristGLM_age_only)
wrists_full= r.squaredGLMM(wristGLM_full)

faces_age_only =  r.squaredGLMM(faceGLM_age_only)
faces_full =r.squaredGLMM(faceGLM_full)

# wrists_no_int= r.squaredGLMM(wristGLM_no_int)
# faces_no_int =  r.squaredGLMM(faceGLM_no_int)
```

These trends were quantified using two generalized linear mixed-effect models estimating the proportion of faces and hands that were in view, with orientation, posture, their interaction, and scaled participant's age as fixed effects, and with random slopes for infants' orientation and posture. A summary of the coefficients of the models can be found in Tables 2 and 3, confirming that infants who were sitting/standing and in front of their caregivers saw the most faces and hands.  When modeling the proportion of faces seen, age was not a significant predictor; however, age remained a significant predictor when modeling the proportion of hands seen by infants.

We directly examined the contributions of posture and orientation by fitting a reduced version of the full model [@nakagawa2013general] without their fixed effects (both models were run with the maximal random effects structure). The fixed effects in a model with only the age of the participants accounted for relatively little variance in the proportion of faces (marginal $R^2$ < `r round(faces_age_only[1,1],5)`) or hands in view  (marginal $R^2$ = `r round(wrists_age_only[1,1],3)`). However, when adding children's posture and orientation to their caregiver to the model (and their interaction) the marginal $R^2$ were higher for both faces (marginal $R^2$ = `r round(faces_full[1,1],3)`) and wrists (marginal $R^2$ = `r round(wrists_full[1,1],3)`).

Overall, these results suggest that infants' visual access to social information is largely modulated by their posture and orientation to their caregiver, which is in turn a function of their general locomotor development. Nonetheless, these results also suggest that infants may still increase their attention to hands throughout this age range as they continue to learn about objects, their functions, and their names.

@@

@@

This dataset differed from the present in two key ways that could present a challenge for our automated methodology. First, the environment that infants were immersed in with their caregivers (and experimenters)  was much larger and more varied than the play room used in the present dataset, containing multiple structures and toys in different parts of the room for infants to explore, unlike the  room used for the present dataset which was relatively small (approximately 10' x 10'). Thus, our automated methodology could fail to generalize to scenes from these more complex environments, where detecting faces and hands could arguably be a much harder task. Second, @franchak2017see used a head-mounted eye-tracker, finding that children's posture affected where children were looking within their visual field. Nonetheless, if children are often orienting their heads towards where they are allocating their attention [@yoshida2008] then we should expect to find the same pattern of results only when analyzing the information in view.
```{r loadFranchakData}
# loads pos_detections used below for franchak dataset
load(file = '../data/franchak_dataset/preprocessed/franchak_dataset_detections.RData')
```

### 2020-04-30 writeup-paper/draft_apa_march2020.tex
edits to paper draft
make figures plausibly work in-line
add marginal r^2 analysis
other typos and edits

Correspondence concerning this article should be addressed to Bria L. Long, 450 Serra Mall, Stanford CA 94305. E-mail: bria@stanford.edu}
\keywords{Postural developments modulate children's visual access to social information\newline\indent Word count: X}
\DeclareDelayedFloatFlavor{ThreePartTable}{table}
\DeclareDelayedFloatFlavor{lltable}{table}
\DeclareDelayedFloatFlavor*{longtable}{table}
\makeatletter
\renewcommand{\efloat@iwrite}[1]{\immediate\expandafter\protected@write\csname efloat@post#1\endcsname{}}
\makeatother
\usepackage{csquotes}

@@

\abstract{
The ability to process social information is a critical component of
children's early linguistic and cognitive development. However, as
children reach their first birthday, they begin to locomote themselves,
dramatically changing the way they see the world around them. How do
these postural and locomotor developments affect children's access to
the social information relevant for word-learning? We explored this
question by using head-mounted cameras to record the egocentric visual
perspective of 36 infants' (at 8, 12, and 16 months of age) during a
naturalistic play session with their caregiver. To estimate the
proportion of faces and hands in the infant view, we used a computer
vision algorithm to detect the presence of faces and hands in the entire
dataset. We found that infants\textbackslash{}`posture and orientation
to their caregiver changed dramatically across this age range and
modulated their access to this social information; infants who were
sitting or standing with their caregiver at a close distance tended to
have the most faces/hands in their visual field. We also applied our
automated analysis to the egocentric video data from a recent paper
documenting how posture changes one-year-olds visual access to faces
(Franchak et al., 2017), finding convergence across both methodologies
and dataset and confirming previous work that motoric developments play
a significant role in the emergence of children's linguistic and social
capacities. We suggest that the combined use of head-mounted cameras and
the application of new computer vision techniques is a promising avenue
for understanding the statistics of infants\textbackslash{}' visual and
linguistic experience as they change over development.
}

@@

\subsubsection{Participants}\label{participants}

@@

\subsubsection{Head-mounted camera}\label{head-mounted-camera}

@@

\subsubsection{Procedure}\label{procedure}

@@

\subsubsection{Data Processing and
Annotation}\label{data-processing-and-annotation}

@@

\begin{figure*}
\centering
\includegraphics[width=5.5in]{images/detector_samples_banner.pdf}
\caption{\label{fig:frames} Example face and pose detections made by OpenPose (top row) and MTCNN (bottom row) from a child in each age group. The last column features a false positive from OpenPose and a false negative from MTCNN.}
\end{figure*}

@@

These trends were quantified using two generalized linear mixed-effect
models estimating the proportion of faces and hands that were in view,
with orientation, posture, their interaction, and scaled participant's
age as fixed effects, and with random slopes for infants' orientation
and posture. A summary of the coefficients of the models can be found in
Table 2, confirming that infants who were sitting/standing and in front
of their caregivers saw the most faces and hands. When modeling the
proportion of faces seen, age was not a significant predictor; however,
age remained a significant predictor when modeling the proportion of
hands seen by infants. Overall, these results suggest that infants'
visual access to social information is largely modulated by their
posture and orientation to their caregiver, which is in turn a function
of their general locomotor development. Nonetheless, these results also
suggest that infants may still increase their attention to hands
throughout this age range as they continue to learn about objects, their
functions, and their names.

@@

>>>>

Correspondence concerning this article should be addressed to Bria L. Long, 450 Serra Mall, Stanford CA 94305. E-mail: bria@stanford.edu}
\keywords{Postural developments modulate children's visual access to social information\newline\indent Word count: 4295}
\DeclareDelayedFloatFlavor{ThreePartTable}{table}
\DeclareDelayedFloatFlavor{lltable}{table}
\DeclareDelayedFloatFlavor*{longtable}{table}
\makeatletter
\renewcommand{\efloat@iwrite}[1]{\immediate\expandafter\protected@write\csname efloat@post#1\endcsname{}}
\makeatother
\usepackage{csquotes}

@@

\abstract{
The ability to process social information is a critical component of
children's early linguistic and cognitive development. However, as
children reach their first birthday, they begin to locomote themselves,
dramatically changing the way they see the world around them. How do
these postural and locomotor developments affect children's access to
the social information relevant for word-learning? We explored this
question by using head-mounted cameras to record the egocentric visual
perspective of 36 infants (at 8, 12, and 16 months of age) during a
naturalistic play session with their caregiver. To estimate the
proportion of faces and hands in the infant view, we used a computer
vision algorithm to detect the presence of faces and hands in the entire
dataset. We found that infants' posture and orientation to their
caregiver changed dramatically across this age range and modulated their
access to this social information; infants who were sitting or standing
with their caregiver at a close distance tended to have the most
faces/hands in their visual field. We also applied our automated
analysis to the egocentric video data from a recent paper documenting
how posture changes one-year-olds visual access to faces (Franchak et
al., 2017), finding convergence across both methodologies and dataset
and confirming previous work that motoric developments play a
significant role in the emergence of children's linguistic and social
capacities. We suggest that the combined use of head-mounted cameras and
the application of new computer vision techniques is a promising avenue
for understanding the statistics of infants' visual and linguistic
experience as they change over development.
}

@@

\subsection{Participants}\label{participants}

@@

\subsection{Head-mounted camera}\label{head-mounted-camera}

@@

\subsection{Procedure}\label{procedure}

@@

\subsection{Data Processing and
Annotation}\label{data-processing-and-annotation}

@@

\begin{figure}[htbp]
\centering
\includegraphics{figs/frames-1.pdf}
\caption{\label{fig:frames}Example face and pose detections made by OpenPose
(top row) and MTCNN (bottom row) from a child in each age group. The
last column features a false positive from OpenPose and a false negative
from MTCNN.}
\end{figure}

@@

These trends were quantified using two generalized linear mixed-effect
models estimating the proportion of faces and hands that were in view,
with orientation, posture, their interaction, and scaled participant's
age as fixed effects, and with random slopes for infants' orientation
and posture. A summary of the coefficients of the models can be found in
Tables 2 and 3, confirming that infants who were sitting/standing and in
front of their caregivers saw the most faces and hands. When modeling
the proportion of faces seen, age was not a significant predictor;
however, age remained a significant predictor when modeling the
proportion of hands seen by infants.

We directly examined the contributions of posture and orientation by
fitting a reduced version of the full model (Nakagawa \& Schielzeth,
2013) without their fixed effects (both models were run with the maximal
random effects structure). The fixed effects in a model with only the
age of the participants accounted for relatively little variance in the
proportion of faces (marginal \(R^2\) = 0.00) or hands in view (marginal
\(R^2\) = 0.15). However, when adding children's posture and orientation
to their caregiver to the model (and their interaction) the marginal
\(R^2\) were higher for both faces (marginal \(R^2\) = 0.28) and wrists
(marginal \(R^2\) = 0.37).

Overall, these results suggest that infants' visual access to social
information is largely modulated by their posture and orientation to
their caregiver, which is in turn a function of their general locomotor
development. Nonetheless, these results also suggest that infants may
still increase their attention to hands throughout this age range as
they continue to learn about objects, their functions, and their names.

@@

\hypertarget{ref-nakagawa2013general}{}
Nakagawa, S., \& Schielzeth, H. (2013). A general and simple method for
obtaining r2 from generalized linear mixed-effects models. \emph{Methods
in Ecology and Evolution}, \emph{4}(2), 133--142.
